---
title: "L12_Analysis"
author: "Weinberg"
date: "2022-11-07"
output: html_document
---
•	Graphs and the results of statistical tests addressing whether or not there is a significant difference in classical conditioning response across life history stage, sex, and individual identity.

Test for effect of the following on learning ability (MLR response to CS+ in retention test):
•	Age
•	Sex
•	Individual variation


#Box 0. Import the data and libraries
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(googlesheets4)
library(ggplot2)

mysheet<-"https://docs.google.com/spreadsheets/d/10dDmL8DCZTpINpvL-wHz_N20N3VQfdEzM3wDK_OOZ44/edit?usp=sharing"
dat<-read_sheet(mysheet)
colnames(dat)<-c("group", "animal", "sex", "instar", "length", "CS_plus", "CS_minus", "control")
dat<-na.omit(dat)
dat
```

#Box 1. View the data and summary statistics
In order to classify an individual as having correctly learned the conditioned stimulus, it must:
-Exhibit the MLR when exposed to the CS+ odor
-Not exhibit MLR when exposed to the CS- or control odor

Print the percentage of cockroaches who learned the stimulus
-By sex 
-By instar

```{r}
dat$result<-0
for(i in 1:nrow(dat)){
  if(dat[i,6]==1 & dat[i,7]==0 & dat[i,8]==0){dat$result[i]<-1}
  else{dat$result[i]<-0}
}
perc<-function(x){sum(x)/length(x)*100}

sumsex<-dat %>% group_by(sex) %>% summarize(percent.learned=perc(result))
suminstar<- dat %>% group_by(instar) %>% summarize(percent.learned=perc(result))
sumsex
suminstar

```

#Box 2a. Plot the data by sex
Replace "my plot title" in the first line below with the title you would like to use for this plot.
```{r}
plot1<-"my plot title"
plot_sex<-ggplot(data = sumsex, aes(x=sex, y=percent.learned, fill=sex))+
  geom_col()+labs(title=plot1)
plot_sex
```

#Box 2b. Plot the data by instar
Replace "my plot title" in the first line below with the title you would like to use for this plot.
```{r}
plot2<-"my plot title"

dat$instar<-as.factor(dat$instar)
plot_instar<-ggplot(data = suminstar, aes(x=instar, y=percent.learned, fill=instar))+
  geom_col()+labs(title=plot2)

plot_instar
```
#Box 2c. Plot individual learning responses 

In this plot, each individual has its own value on the x-axis. A value of "1" means that the individual learned the behavior, while a value of "0" means that it did not.

Replace "my plot title" in the first line below with the title you would like to use for this plot.

```{r}
plot3<-"my plot title"

plot_instar<-ggplot(data = dat %>% mutate(ID=row_number()), aes(x=ID, y=result))+
  geom_point(aes(shape=instar, color=sex))+labs(title=plot3)

plot_instar
```

#Box 3. Run statistical tests (ANOVA) to investigate whether differences in learning ability vary significantly with:
-age
-sex
-interaction between age and sex

The value Pr(>F) is the p-value for this test.
```{r}
dat$sex<-as.factor(dat$sex)
sex.aov<-aov(result~sex, data=dat)

instar.aov<-aov(result~instar, data=dat)

intx.aov<-aov(result~sex*instar, data=dat)

summary(instar.aov)
summary(sex.aov)
summary(intx.aov)

```
