---
title: "L1_Graphs"
author: "Weinberg"
date: "2022-08-13"
output: html_document
---
Step 0. Load the packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
install.packages(c("tidyverse", "ggplot2", "googlesheets4"))
library(tidyverse)
library(ggplot2)
library(googlesheets4)
```


Step 1. Import the data
Make sure the spreadsheet has only one tab with your data (do not copy the entire example spreadsheet) 
Import your data by replacing "https://docs.google.com/myspreadhseet" with the actual sharing link to your data sheet
You may be prompted to log in to your UC Berkeley account
```{r readdat}
mysheet<-"https://docs.google.com/spreadsheets/d/1hxgHexndeOuiQKPYr9OFIluTmtNi0Zrwiu0noeopf88/edit?usp=sharing"
```


####PART 1: ONLY RUN IF YOU USED FREQUENCY SAMPLING####
Replace "condition 1" and "condition 2" below with the names of the columns in your data table
```{r name frequency}
datf<-read_sheet(mysheet)
colnames(datf)<-c("behavior", "condition 1", "condition 2")
```

We'll keep things simple for the first lab. The output of the plot function below is the only thing you need to turn in with the lab assignment.
```{r plot frequency}
pf<-ggplot(data=datf %>% pivot_longer(-behavior), aes(x=behavior, y=value)) +
 geom_col(aes(fill=name), position="dodge")

pf
```
When you visually inspect the graph, do you see any differences between the two conditions (or individuals) you tested?


###PART 2: ONLY RUN IF YOU USED INTERVAL SAMPLING####

Step 1. Import the data
Make sure the spreadsheet has only one tab with your data (do not copy the entire example spreadsheet) 
Import your data by replacing "https://docs.google.com/myspreadhseet" with the actual sharing link to your data sheet
You may be prompted to log in to your UC Berkeley account
```{r import interval}
mysheet<-"https://docs.google.com/spreadsheets/d/1UEcyCuOiHbV57b1EUDUwIapW7D8f48LYYDXro3O3b88/edit?usp=sharing"
dati<-read_sheet(mysheet)
dati<-dati[,2:ncol(dati)] #this removes the column with time points. In this lab,  we are interested in looking at overall occurrences of a behavior but we are not concerned about when they occurred 
```


Replace "behavior1", etc. with the names of the behaviors  in your study. Add or remove entries as needed so that you have the correct number. If you add an entry, make sure it is in quotes and separated from the other entries with a comma.
```{r name interval}
colnames(dati)<-c("condition", "behavior1", "behavior2", "behavior3", "behavior4")

```


```{r plot interval}
pi<-ggplot(data=dati %>% pivot_longer(cols = 2:ncol(dati), names_to="behavior") %>% group_by(condition, behavior) %>% summarise(count=sum(value)), aes(x=behavior, y=count)) +
 geom_col(aes(fill=condition), position="dodge")

pi
```


