---
title: "L4_analysis"
author: "Weinberg"
date: "2022-09-13"
output: html_document
---


#Box 0. Run to load the libraries
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(googlesheets4)
library(ggplot2)
```


#Box 1. Import the data
This is a link to the class data. Do NOT change the link.
```{r}
mysheet<-"https://docs.google.com/spreadsheets/d/1reC0m_6hRsa9UCoGNrBf1JAzeWMrdBGBflWBr5gkD38/edit?usp=sharing"
pt1<-read_sheet(mysheet, sheet="Part 1")
pt2<-read_sheet(mysheet, sheet="Part 2")
pt3<-read_sheet(mysheet, sheet="Part 3")

conditions1<-c("unmarked.no.visual", "marked.no.visual", "unmarked.visual", "marked.visual")
colnames(pt1)<-c("nest", conditions1)

conditions2<-c("unmarked.chemical.only", "marked.chemical.only", "unmarked.visual.only", "marked.visual.only")
colnames(pt2)<-c("nest",conditions2)


colnames(pt3)<-c("nest", "angle", "to_nest", "to_platform")
```


###PART 1###

#Box 2. Visualize data from part 1
First plot the number of marked ants (those that had the opportunity to learn a visual cue) crossing over to platform #1 (no cue) and platform #2 (with cue)

```{r}
plot_pt1<-ggplot(pt1 %>% pivot_longer(cols=c(conditions1), names_to="condition", values_to="count"), aes(x=condition, y=count, fill=condition))+
  geom_boxplot()+
  ggtitle("Trail preference in ants with and without prior exposure to visual cue") +theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

plot_pt1
```

#Box 3. Descriptive statistics for part 1
In your lab report, write the mean and standard deviation of number of ants in each group observed traveling to each platform (four means and four standard deviations total)
```{r}
sum_1<- pt1 %>% pivot_longer(cols=conditions1, names_to="condition", values_to="count") %>% group_by(condition) %>% summarise(mean=mean(count), stdev=sd(count))

sum_1
```


#Box 4. Statistical tests for difference in bridge usage in exposed (marked) ants.

The t-tests performed here will address the following questions (respectively)
1) Did ants that had the opportunity to learn a visual cue preferentially forage on the bridge with the cue?
2) Did unmarked ants show a preference for either bridge?

The first result printed below is the t-test for bridge preference in exposed (marked) ants. In this test, x=marked ants crossing the bridge with no visual cue, y=marked ants crossing the bridge with the visual cue

The second result is the t-test for unexposed (unmarked) ants. 
In this test, x=unmarked ants crossing the bridge with no visual cue, y=unmarked ants crossing the bridge with a visual cue

Note that we expect the presence or absence of a "visual cue" to be irrelevant for these ants since they had no prior exposure to the visual cue. What would it indicate if there was a significant difference in bridge usage for unmarked ants?


```{r}
marked_test1<-t.test(x=pt1$marked.no.visual, y=pt1$marked.visual)

unmarked_test1<-t.test(x=pt1$unmarked.no.visual, y=pt1$unmarked.visual)

marked_test1
unmarked_test1

```



###PART 2###

#Box 5. Plot results for part 2
Note that "marked" ants now includes both green and yellow ants, both of which were exposed to the visual cue, either in part 1 (green) or part 2 (yellow).

Remember, platform 1 had a chemical cue only while platform 2 had a visual cue only.

```{r}
plot_pt2<-ggplot(pt2 %>% pivot_longer(cols=c(conditions2), names_to="condition", values_to="count"), aes(x=condition, y=count, fill=condition))+
  geom_boxplot()+
  ggtitle("Trail usage by ants with and without prior exposure to visual cue")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))


plot_pt2

```

#Box 6. Descriptive statistics for part 2

Next we will look at the mean and standard deviation of ants on the trail with the visual cue only and the chemical cue only

```{r}
sum_2<- pt2 %>% pivot_longer(cols=conditions2, names_to="condition", values_to="count") %>% group_by(condition) %>% summarise(mean=mean(count), stdev=sd(count))

sum_2
```

#Box 7. Statistical tests for part 2

Here we will again use a t-test to test for differences in bridge usage for both (1) marked and (2) unmarked ants 
These tests will address the following questions (respectively)
1) Did ants with prior exposure to a visual cue show a preference for the bridge with the visual, but no chemical, cue or the bridge with a chemical, but no visual, cue?

2)Did ants with no prior exposure to a visual cue show a preference for the bridge without the chemical cue (visual.only) or with the chemical cue (chemical.only)?

```{r}
marked_test2<-t.test(x=pt2$marked.visual.only, y=pt2$marked.chemical.only)

unmarked_test2<-t.test(x=pt2$unmarked.visual.only, y=pt2$unmarked.chemical.only)

marked_test2
unmarked_test2
```

###PART 3###

#Box 8. Plot correct (towards nest) vs. incorrect (towards opposite platform) turns made by ants on trails with different bifurcation angles

```{r}

plot_pt3<- ggplot(pt3 %>% pivot_longer(cols=c(to_nest, to_platform), names_to="direction", values_to="count") %>% group_by(angle), aes(x=direction, y=count, fill=as.factor(angle))) +
  geom_boxplot() + labs(fill="bifurcation angle", title="Orientation of Argentine ants using different trail geometries")

plot_pt3
```


#Box 9. Descriptive statistics for trail geometry

```{r}
sum_3<- pt3 %>% pivot_longer(cols=c(to_nest, to_platform), names_to="direction", values_to="count") %>% group_by(angle, direction) %>% summarise(mean=mean(count), stdev=sd(count))

sum_3

```


#Box 10. 
This box will use a t-test to compare the percentage of ants taking the correct path (towards the nest) 
In the results, x represents the percentage of ants taking the correct path when the bifurcatin was 60 degrees and y represents the percentage of ants taking the correct path when the bifurcation was 120 degrees.
```{r}
perc_by_angle<-pt3 %>% group_by(angle) %>% mutate(perc_correct=(to_nest/sum(to_nest, to_platform))*100)

geotest<-t.test(x=perc_by_angle %>% filter(angle==60), y=perc_by_angle %>% filter(angle==120))

geotest
```

#Box 11. Plot the percentage of ants taking the correct turn by angle
```{r}
plot_perc<-ggplot(data=perc_by_angle, aes(x=angle, y=perc_correct, group=angle, fill=angle))+
  geom_boxplot() +theme(legend.position = "none")

plot_perc
```
